An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-12-18 jmiv.tex
h=0.25 figures; dashed lines in plots

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods. \revision{Moreover, parametric approaches to Elastica  as snakes and splines depend on the choice of anchor points and  multigrid convergence hasn't been proven for them.}

@@

\begin{figure}[!h]
\center
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h1/summary.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h05/summary.eps}
	}%
	\hspace{10pt}
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h1/summary.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h05/summary.eps}
	}%
		\caption{\revision{Local combinatorial optimization process results using II estimator with radius $7$ for the square and flower shapes. Shapes displayed at every $2$ iterations for the square and $5$ iterations for the flower.}}	
		\label{fig:local-comb-square-results}
\end{figure}

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a \revision{$32$-core $2.4Ghz$ CPU and the number of pixels in the square and
flower shapes are respectively $841,1867$ for grid step $h=1.0$}. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. \revision{In particular, the digital elastica energy for the experiments in figure \ref{fig:local-comb-square-results} approaches the global optimum (figure  \ref{fig:plot-elastica-local-search})}. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore, we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

\begin{figure}[!h]
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Digital elastica computation along iterations of algorithm \ref{alg:local-search}. \revision{The final shape has energy value closer to the optimal value $2\pi/5$ (dashed line)}.}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\revision{
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 4s (0.4s/it) & 77s (3s/it)\\
\hline
Flower & 252s (4s/it)  & 3246s (42s/it)\\
\hline
\end{tabular}
}
\caption{Running times for the local combinatorial optimization algorithm with $\revision{ 0 \leq n \leq 50}$. Thirty two threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

\begin{algorithm}[H]
 \SetKwData{It}{i}
 \SetKwData{MIt}{maxIt}
 \SetKwData{Delta}{delta}
 \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
 \SetKwComment{comment}{//}{}

@@

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}\\
		\subfloat[\revision{Digital elastica evaluation}]{%
		\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/square-flower-triangle.eps}						
		}
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). In figure (c), we observe that for a given choice of radius, the digital elastica may increase after a certain number of iterations. }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}

@@

\begin{figure}[h!]
\center
\includegraphics[scale=0.5]{images/optimization/pairwise-ratio/plot-pairwiseratio-lowerHigher-concavities-probe.eps}
\caption{We plot the ratio of pairwise terms among all $\binom{|X^{(i)}|}{2}$ combinations. The highest ring has roughly half the number of pairwise terms as the lowest ring.}
\label{fig:ratio-pairwise-terms}
\end{figure}

@@

\begin{algorithm}[H]
 \SetKwData{It}{i}
 \SetKwData{MIt}{maxIt}
 \SetKwData{Tol}{tolerance}
 \SetKwData{Delta}{delta}
 \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
 \SetKwComment{comment}{//}{}

>>>>

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods. \revision{Moreover, parametric approaches to Elastica  as snakes and splines are way to sensitive to the choice of anchor points and  multigrid convergence hasn't been proven for them.}

@@

\begin{figure}[!h]
	\center
	\begin{tabular}{ccc}
		$h=1.0$ & $h=0.5$ & $h=0.25$\\
		\includegraphics[scale=0.2]{images/local_search/square/h1/summary.eps} & \includegraphics[scale=0.2]{images/local_search/square/h05/summary.eps} & \includegraphics[scale=0.2]{images/local_search/square/h025/summary.eps} \\
		\includegraphics[scale=0.2]{images/local_search/flower/h1/summary.eps} & \includegraphics[scale=0.2]{images/local_search/flower/h05/summary.eps} & \includegraphics[scale=0.2]{images/local_search/flower/h025/summary.eps}
	\end{tabular}	
		\caption{\revision{Local combinatorial optimization process results using II estimator with radius $5$ for the square and flower shapes. Shapes are displayed at every $5$ iterations.}}	
		\label{fig:local-comb-square-results}		
\end{figure}

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a \revision{$32$-core $2.4Ghz$ CPU and the number of pixels in the square and
flower shapes are respectively $841,1641$ for grid step $h=1.0$}. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. \revision{In particular, the digital elastica energy for the experiments in figure \ref{fig:local-comb-square-results} approaches the global optimum (figure  \ref{fig:plot-elastica-local-search})}. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore, we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

\begin{figure}
	\center
	\captionsetup{type=table}	
	\revision{
	\begin{tabular}{|c|c|c|}
	\hline
	& Square & Flower \\
	\hline
	$h=1.0$ & 4s & 124s \\
	\hline
	$h=0.5$ & 77s & 1924s\\
	\hline
	$h=0.25$ & 1886s & 28445s\\
	\hline
	\end{tabular}
	}
	\caption{Running times for the local combinatorial optimization algorithm with $0 \leq n \leq 50$. Thirty two threads were used.}
	\label{tab:summary-local-comb-rtime}
\end{figure}

\begin{figure}[!h]
\center
\subfloat[\label{}]{\includegraphics[scale=0.6]{images/local_search/square.eps}}
\subfloat[\label{}]{\includegraphics[scale=0.6]{images/local_search/flower.eps}}
\caption{Digital elastica computation along iterations of algorithm \ref{alg:local-search}. \revision{The final shape has energy value closer to the optimal value $2\pi/5$ (thin horizontal dashed line)}.}
\label{fig:plot-elastica-local-search}
\end{figure}

@@

\begin{algorithm}
 \SetKwData{It}{i}
 \SetKwData{MIt}{maxIt}
 \SetKwData{Delta}{delta}
 \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
 \SetKwComment{comment}{//}{}

@@

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}\\
		\subfloat[\revision{Digital elastica evaluation ($r=5,h=0.5$)}]{%
		\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/square-flower-triangle.eps}						
		}
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). In figure (c), we observe that for a given choice of radius, the digital elastica may increase after a certain number of iterations. }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[scale=0.5]{images/optimization/pairwise-ratio/plot-pairwiseratio-lowerHigher-concavities-probe.eps}
\caption{We plot the ratio of pairwise terms among all $\binom{|X^{(i)}|}{2}$ combinations. The highest ring has roughly half the number of pairwise terms as the lowest ring.}
\label{fig:ratio-pairwise-terms}
\end{figure}

@@

\begin{algorithm}
 \SetKwData{It}{i}
 \SetKwData{MIt}{maxIt}
 \SetKwData{Tol}{tolerance}
 \SetKwData{Delta}{delta}
 \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
 \SetKwComment{comment}{//}{}
